>

### 2019-04-12 paper/spinifex-paper.tex
Section level changes. DOI's to ref. building out code/package section.

% !TeX root = RJwrapper.tex
\title{spinifex: manual control of dynamic linear projections of
high-dimensional data}
\author{by Nicholas Spyrison, Dianne Cook}

@@

\bibliography{RJrefs}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

@@

\hypertarget{sec:algorithm}{%
\subsection{Algorithm}\label{sec:algorithm}}

@@

\hypertarget{notation}{%
\subsubsection{Notation}\label{notation}}

@@

\hypertarget{toy-data-set}{%
\subsubsection{Toy data set}\label{toy-data-set}}

@@

\begin{Schunk}
\begin{Sinput}
# devtools::install_github("nspyrison/spinifex") # Development version
install.package("spinifex")
library("spinifex")

# Also see the vignette: 
vignette("spinifex")

@@

\hypertarget{step-1-choose-variable-of-interest}{%
\subsubsection{Step 1) Choose variable of
interest}\label{step-1-choose-variable-of-interest}}

@@

\hypertarget{step-2-create-the-manip-space}{%
\subsubsection{Step 2) Create the manip
space}\label{step-2-create-the-manip-space}}

@@

\begin{align*}
  \textbf{e}_k &\leftarrow 1 \\ 
  \textbf{e}   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
  &\\
  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
\end{align*}

@@

\hypertarget{step-3-generate-rotation}{%
\subsubsection{Step 3) Generate
rotation}\label{step-3-generate-rotation}}

@@

\textbf{For } \(i\) \textbf{in 1 to n\_slides:}

Post-multiply the manipulation space by the pre-defined rotation matrix
producing \textbf{RM}, the rotated manip space.

@@

then \begin{align*}
  \textbf{RM}_{[p,~3,~i]}
  &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]} \\
  &= \textbf{M}_{[p,~3]}
    ~*~
  \begin{bmatrix}
    c_\theta^2 c_\phi s_\theta^2 &
    -c_\theta s_\theta (1 - c_\phi) &
    -c_\theta s_\phi \\
    -c_\theta s_\theta (1 - c_\phi) &
    s_\theta^2 c_\phi + c_\theta^2 &
    -s_\theta s_\phi \\
    c_\theta s_\phi &
    s_\theta s_\phi &
    c_\phi
  \end{bmatrix}_{[3,~3,~i]}
\end{align*}

@@

\hypertarget{sec:display}{%
\subsection{Data in projection-space}\label{sec:display}}

@@

\hypertarget{rendering-and-sharing}{%
\subsubsection{Rendering and sharing}\label{rendering-and-sharing}}

@@

\hypertarget{storage}{%
\subsubsection{Storage}\label{storage}}

@@

\hypertarget{sec:application}{%
\subsection{Application}\label{sec:application}}

@@

\hypertarget{jet-cluster}{%
\subsubsection{Jet cluster}\label{jet-cluster}}

@@

\hypertarget{dis-cluster}{%
\subsubsection{DIS cluster}\label{dis-cluster}}

@@

\hypertarget{summary}{%
\subsubsection{Summary}\label{summary}}

@@

\hypertarget{sec:discussion}{%
\subsection{Discussion}\label{sec:discussion}}

@@

\hypertarget{acknowledgments}{%
\subsection{Acknowledgments}\label{acknowledgments}}

@@

\hypertarget{bibliography}{%
\subsection*{Bibliography}\label{bibliography}}
\addcontentsline{toc}{subsection}{Bibliography}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-asimov_grand_1985}{}%
Asimov, Daniel. 1985. ``The Grand Tour: A Tool for Viewing
Multidimensional Data.'' \emph{SIAM Journal on Scientific and
Statistical Computing} 6 (1): 128--43.

@@

\leavevmode\hypertarget{ref-cook_dynamical_2018}{}%
Cook, Dianne, Ursula Laa, and German Valencia. 2018. ``Dynamical
Projections for the Visualization of PDFSense Data.'' \emph{Eur. Phys.
J. C} 78 (9): 742.

@@

\leavevmode\hypertarget{ref-kirkpatrick_optimization_1983}{}%
Kirkpatrick, Scott, C. Daniel Gelatt, and Mario P. Vecchi. 1983.
``Optimization by Simulated Annealing.'' \emph{Science} 220 (4598):
671--80.

\leavevmode\hypertarget{ref-lubischew_use_1962}{}%
Lubischew, Alexander A. 1962. ``On the Use of Discriminant Functions in
Taxonomy.'' \emph{Biometrics}, 455--77.

@@

\leavevmode\hypertarget{ref-wang_mapping_2018}{}%
Wang, Bo-Ting, T. J. Hobbs, Sean Doyle, Jun Gao, Tie-Jiun Hou, Pavel M.
Nadolsky, and Fredrick I. Olness. 2018. ``Mapping the Sensitivity of
Hadronic Experiments to Nucleon Structure.'' \emph{Physical Review D} 98
(9): 094030.

@@

\bibliography{RJrefs.bib}

>>>>

% !TeX root = RJwrapper.tex
\title{spinifex: A Package for Manual Control of Dynamic Linear Projections of
Multivariate Data}
\author{by Nicholas Spyrison, Dianne Cook}

@@

\bibliography{spinifex_paper}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\hypertarget{sec:algorithm}{%
\section{Algorithm}\label{sec:algorithm}}

@@

\hypertarget{notation}{%
\subsection{Notation}\label{notation}}

@@

\hypertarget{toy-data-set}{%
\subsection{Toy data set}\label{toy-data-set}}

@@

\begin{Schunk}
\begin{Sinput}
# remotes::install_github("nspyrison/spinifex") # Development version
install.package("spinifex")
library("spinifex")

# Also see the vignette: 
vignette("spinifex_vignette")

@@

\hypertarget{step-1-choose-variable-of-interest}{%
\subsection{Step 1) Choose variable of
interest}\label{step-1-choose-variable-of-interest}}

@@

\hypertarget{step-2-create-the-manip-space}{%
\subsection{Step 2) Create the manip
space}\label{step-2-create-the-manip-space}}

@@

\begin{align*}
  \textbf{e}_k &\leftarrow 1 \\ 
  \textbf{e}   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
\end{align*}

@@

\hypertarget{step-3-generate-rotation}{%
\subsection{Step 3) Generate rotation}\label{step-3-generate-rotation}}

@@

Post-multiply the manipulation space by the pre-defined rotation matrix
producing \textbf{RM}, the rotated manip space. This is one frame of
basis values, repeat this procress for each value in the sequences of
\(\theta\) and \(\phi\) for the complete animation.

@@

\textbf{For each value of } \(i\) \textbf{:}

\begin{align*}
  \textbf{RM}_{[p,~3,~i]}
  &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]} \\
  &= \textbf{M}_{[p,~3]}
    ~*~
  \begin{bmatrix}
    c_\theta^2 c_\phi s_\theta^2 &
    -c_\theta s_\theta (1 - c_\phi) &
    -c_\theta s_\phi \\
    -c_\theta s_\theta (1 - c_\phi) &
    s_\theta^2 c_\phi + c_\theta^2 &
    -s_\theta s_\phi \\
    c_\theta s_\phi &
    s_\theta s_\phi &
    c_\phi
  \end{bmatrix}_{[3,~3,~i]}
\end{align*}

@@

\hypertarget{sec:display}{%
\section{Data in projection-space}\label{sec:display}}

@@

\hypertarget{package-structure-and-functionality}{%
\section{Package structure and
functionality}\label{package-structure-and-functionality}}

\hypertarget{installation}{%
\subsection{Installation}\label{installation}}

\hypertarget{walk-though-package-functions-in-terms-of-work-flow.}{%
\subsection{Walk though package functions in terms of work
flow.}\label{walk-though-package-functions-in-terms-of-work-flow.}}

\hypertarget{rendering-and-sharing}{%
\subsection{Rendering and sharing}\label{rendering-and-sharing}}

@@

\hypertarget{storage}{%
\subsection{Storage}\label{storage}}

@@

\hypertarget{sec:application}{%
\section{Application}\label{sec:application}}

@@

\hypertarget{jet-cluster}{%
\subsection{Jet cluster}\label{jet-cluster}}

@@

\hypertarget{dis-cluster}{%
\subsection{DIS cluster}\label{dis-cluster}}

@@

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

@@

\hypertarget{sec:discussion}{%
\section{Discussion}\label{sec:discussion}}

@@

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

@@

\hypertarget{bibliography}{%
\section*{Bibliography}\label{bibliography}}
\addcontentsline{toc}{section}{Bibliography}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-asimov_grand_1985}{}%
Asimov, Daniel. 1985. ``The Grand Tour: A Tool for Viewing
Multidimensional Data.'' \emph{SIAM Journal on Scientific and
Statistical Computing} 6 (1): 128--43.
\url{https://doi.org/https://doi.org/10.1137/0906011}.

@@

\leavevmode\hypertarget{ref-cook_dynamical_2018}{}%
Cook, Dianne, Ursula Laa, and German Valencia. 2018. ``Dynamical
Projections for the Visualization of PDFSense Data.'' \emph{Eur. Phys.
J. C} 78 (9): 742. \url{https://doi.org/10.1140/epjc/s10052-018-6205-2}.

@@

\leavevmode\hypertarget{ref-kirkpatrick_optimization_1983}{}%
Kirkpatrick, Scott, C. Daniel Gelatt, and Mario P. Vecchi. 1983.
``Optimization by Simulated Annealing.'' \emph{Science} 220 (4598):
671--80. \url{https://doi.org/10.1126/science.220.4598.671}.

\leavevmode\hypertarget{ref-lubischew_use_1962}{}%
Lubischew, Alexander A. 1962. ``On the Use of Discriminant Functions in
Taxonomy.'' \emph{Biometrics}, 455--77.
\url{https://doi.org/10.2307/2527894}.

@@

\leavevmode\hypertarget{ref-wang_mapping_2018}{}%
Wang, Bo-Ting, T. J. Hobbs, Sean Doyle, Jun Gao, Tie-Jiun Hou, Pavel M.
Nadolsky, and Fredrick I. Olness. 2018. ``Mapping the Sensitivity of
Hadronic Experiments to Nucleon Structure.'' \emph{Physical Review D} 98
(9): 094030. \url{https://doi.org/10.1103/PhysRevD.98.094030}.

@@

\bibliography{spinifex\_paper.bib}

### 2019-04-12 paper/spinifex_paper.Rmd
Section level changes. DOI's to ref. building out code/package section.

---
title: > 
  spinifex: manual control of dynamic linear projections of high-dimensional data
author:
  - name: Nicholas Spyrison
    affiliation: Monash University 
    address:
    - Faculty of Information Technology
    email:  Nicholas.Spyrison@monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address:
    - Department of Econometrics and Business Statistics
    email:  DiCook@monash.edu
abstract: >
  The set of dynamic linear projections of multivariate data collectively known as **tours** provides an important tool for extending the dimensionality of visuals. The R package **tourr** offers a variety of path generators and geometric displays for conducting tours. This paper discusses an extension package, **spinifex**, that adds support for the path generation of manual tours and extends the display of tours to use with the animation packages, **plotly** and **gganimate**. Manual tours are used to explore the sensitivity of structure as the contributions of a manipulation variable are changed.

  A recent paper [@wang_mapping_2018] visualized the sensitivity of the hadronic experiments to nucleon structure. Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components. This research applies manual tours to this data showing that manual tours resolve information about the structure of cluster distribution by bringing to light information that is orthogonal to static viewing planes.
keywords: manual tour, guided tour, grand tour, projection pursuit, high dimensional data, multivariate data, data visualization, statistical graphics, data science
preamble: >
  % preamble packages
link-citations: true
fig_caption: yes
toc: false
bibliography: RJrefs.bib
output: rticles::rjournal_article
  #bookdown::pdf_book:
   # base_format: rticles::rjournal_article #see sandbox for hard code solution.
---
\bibliography{RJrefs}
<!-- note that Di/Ursula have different setup and options and use static images.  -->
```{r setup_paper, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center", echo = TRUE, collapse = TRUE,
  message = FALSE, warning = FALSE, error = FALSE,
  cache = F, cache.lazy = FALSE
)
library(spinifex)
library(ggplot2)
```

@@

## Introduction

@@

<!-- Paper outline -->
The paper is organized as follows. Section \ref{sec:algorithm} explains the algorithm using a toy dataset. Section \ref{sec:display} discusses the display of the animation after the path has been generated. Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to contemporary high energy physics. The last section, \ref{sec:discussion} summarizes the work and discusses future research.

<!-- Algorithm outline -->
## Algorithm {#sec:algorithm}

@@

The steps are described in more detail below with description, equations, figures, and their code usage.

### Notation

@@

### Toy data set

@@

<!-- step 0 disp code. -->
```{r step0CODEEXAMPLE, eval=F, echo=T}
# devtools::install_github("nspyrison/spinifex") # Development version
install.package("spinifex")
library("spinifex")

# Also see the vignette: 
vignette("spinifex")

f_data  <- tourr::rescale(flea[, 1:6])
f_path  <- save_history(f_data, guided_tour(holes()))
f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
f_cat   <- factor(flea$species)

### Step 1) Choose variable of interest

@@

### Step 2) Create the manip space

@@

<!-- step 2 disp code. -->
```{r step2CODEEXAMPLE, eval=F, echo=T}
view_manip_space(basis = f_basis, 
                 manip_var = 5, 
                 labels = colnames(f_data))
```

### Step 3) Generate rotation

@@

**For ** $i$ **in 1 to n_slides:**
<!-- don't mix rmd and latex... -->

Post-multiply the manipulation space by the pre-defined rotation matrix producing **RM**, the rotated manip space.

@@

then
\begin{align*}
  \textbf{RM}_{[p,~3,~i]}
  &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]} \\
  &= \textbf{M}_{[p,~3]}
    ~*~
  \begin{bmatrix}
    c_\theta^2 c_\phi s_\theta^2 &
    -c_\theta s_\theta (1 - c_\phi) &
    -c_\theta s_\phi \\
    -c_\theta s_\theta (1 - c_\phi) &
    s_\theta^2 c_\phi + c_\theta^2 &
    -s_\theta s_\phi \\
    c_\theta s_\phi &
    s_\theta s_\phi &
    c_\phi
  \end{bmatrix}_{[3,~3,~i]}
\end{align*}

@@

play_manual_tour(data = f_data,
                 basis = f_basis, 
                 manip_var = 5, 
                 angle = angle_speed,
                 col = f_cat,
                 pch = f_cat)
```

## Data in projection-space {#sec:display}

@@

### Rendering and sharing

@@

### Storage

@@

## Application {#sec:application}

@@

### Jet cluster

@@

### DIS cluster

@@

### Summary

@@

## Discussion {#sec:discussion}

@@

## Acknowledgments

@@

## Bibliography

>>>>

---
title: > 
  spinifex: A Package for Manual Control of Dynamic Linear Projections of Multivariate Data
author:
  - name: Nicholas Spyrison
    affiliation: Monash University 
    address:
    - Faculty of Information Technology
    email:  Nicholas.Spyrison@monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address:
    - Department of Econometrics and Business Statistics
    email:  DiCook@monash.edu
abstract: >
  The set of dynamic linear projections of multivariate data collectively known as **tours** provides an important tool for extending the dimensionality of visuals. The R package **tourr** offers a variety of path generators and geometric displays for conducting tours. This paper discusses an extension package, **spinifex**, that adds support for the path generation of manual tours and extends the display of tours to use with the animation packages, **plotly** and **gganimate**. Manual tours are used to explore the sensitivity of structure as the contributions of a manipulation variable are changed.

  A recent paper [@wang_mapping_2018] visualized the sensitivity of the hadronic experiments to nucleon structure. Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components. This research applies manual tours to this data showing that manual tours resolve information about the structure of cluster distribution by bringing to light information that is orthogonal to static viewing planes.
keywords: manual tour, guided tour, grand tour, projection pursuit, high dimensional data, multivariate data, data visualization, statistical graphics, data science
preamble: >
  % preamble packages
link-citations: true
fig_caption: yes
toc: false
bibliography: spinifex_paper.bib
output: rticles::rjournal_article
  #bookdown::pdf_book:
   # base_format: rticles::rjournal_article #see sandbox for hard code solution.
---
\bibliography{spinifex_paper}
<!-- note that Di/Ursula have different setup and options and use static images.  -->
```{r setup_paper, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center", echo = TRUE, collapse = TRUE,
  message = FALSE, warning = FALSE, error = FALSE,
  cache = F, cache.lazy = FALSE
)
library(spinifex)
library(ggplot2)
```

@@

# Introduction

@@

<!-- Paper outline -->
The paper is organized as follows. Section \ref{sec:algorithm} explains the algorithm using a toy dataset. Section \ref{sec:display} discusses the display of the animation after the path has been generated. Section \ref{ref:usage} explains the though the code used in the algorithm section and discusses function structure for the rest of the package. Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to contemporary high energy physics. The last section, \ref{sec:discussion} summarizes the work and discusses future research.

<!-- Algorithm outline -->
# Algorithm {#sec:algorithm}

@@

The steps are described in more detail below with description, equations, and figures. Function names are mentioned inline, but the code is aggregated in

## Notation

@@

## Toy data set

@@

## Step 1) Choose variable of interest

@@

## Step 2) Create the manip space

@@

## Step 3) Generate rotation

@@

Post-multiply the manipulation space by the pre-defined rotation matrix producing **RM**, the rotated manip space. This is one frame of basis values, repeat this procress for each value in the sequences of $\theta$ and $\phi$ for the complete animation.

@@

**For each value of ** $i$ **:**
<!-- don't mix rmd and latex... -->

\begin{align*}
  \textbf{RM}_{[p,~3,~i]}
  &= \textbf{M}_{[p,~3]} ~*~ \textbf{R}_{[3,~3,~i]} \\
  &= \textbf{M}_{[p,~3]}
    ~*~
  \begin{bmatrix}
    c_\theta^2 c_\phi s_\theta^2 &
    -c_\theta s_\theta (1 - c_\phi) &
    -c_\theta s_\phi \\
    -c_\theta s_\theta (1 - c_\phi) &
    s_\theta^2 c_\phi + c_\theta^2 &
    -s_\theta s_\phi \\
    c_\theta s_\phi &
    s_\theta s_\phi &
    c_\phi
  \end{bmatrix}_{[3,~3,~i]}
\end{align*}

@@

# Data in projection-space {#sec:display}

@@

# Package structure and functionality {#sec:usage}

## Installation

<!-- step 0 disp code. -->
```{r installCODEEXAMPLE, eval=F, echo=T}
# remotes::install_github("nspyrison/spinifex") # Development version
install.package("spinifex")
library("spinifex")

# Also see the vignette: 
vignette("spinifex_vignette")
```

## Algorithm

We'll start by initializing values including: standardizing our data set (numeric coulmns only), provide a basis, a categorical variable for point asthetics, and specifiy the manip var. Then we the references axes of the basis with `view_basis()`.

<!-- setup -->
```{r step0CODEEXAMPLE, eval=F, echo=T}
f_data  <- tourr::rescale(flea[, 1:6])                    ## standardize data
f_path  <- save_history(f_data, guided_tour(holes()))     ## produce guided tour
f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2) ## end of guided tour
f_cat   <- factor(flea$species)                           ## categorical var

After becoming familar with the space, select a

<!-- step 2 disp code. -->
```{r step2CODEEXAMPLE, eval=F, echo=T}
f_mvar  <- 5  ## manip var numebr

view_manip_space(basis = f_basis, 
                 manip_var = f_mvar, 
                 labels = colnames(f_data))
```

play_manual_tour(data = f_data,
                 basis = f_basis, 
                 manip_var = f_mvar, 
                 angle = angle_speed,
                 col = f_cat,
                 pch = f_cat)
```

## Walk though package functions in terms of work flow.

## Rendering and sharing

@@

## Storage

@@

# Application {#sec:application}

@@

## Jet cluster

@@

## DIS cluster

@@

## Summary

@@

# Discussion {#sec:discussion}

@@

# Acknowledgments

@@

# Bibliography
